\documentclass[runningheads]{llncs}

% ===== Packages =====
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,backgrounds}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[hidelinks]{hyperref}

% ===== Math shortcuts =====
\newcommand{\kk}{\Bbbk}                 % the base field
\newcommand{\kQ}{\kk Q}                 % path algebra
\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\Adj}{A}                    % adjacency matrix
\newcommand{\walk}{\rightsquigarrow}
\DeclareMathOperator{\softmax}{softmax}
\DeclareMathOperator{\supp}{supp}

\begin{document}

\title{Path Algebras of Quivers as a Sparse\\ Multi-Hop Attention for Mitigating\\ Over-Squashing in GNNs}
\titlerunning{Walk Attention}

\author{Carlos Isaac Zainea Maya\inst{1}\orcidID{0000-0002-6459-2397} \and
Agust\'in Moreno Ca\~nadas\inst{1}\orcidID{0000-0001-6812-5131}}
\authorrunning{C. I. Zainea Maya and A. Moreno Ca\~nadas}
\institute{Universidad Nacional de Colombia, Departamento de Matem\'aticas,\\
Bogot\'a, Colombia\\
\email{\{cizaineam,amorenoca\}@unal.edu.co}}

\maketitle

\begin{abstract}
Message-passing graph neural networks (GNNs) suffer from \emph{over-squashing}:
when information must travel along many long paths into a node, it is compressed
into a fixed-width vector and the signal is lost. We introduce \textbf{Walk
Attention}, an architecture that addresses over-squashing through the
\emph{path algebra} $\kQ$ of the directed graph (quiver) underlying the data,
whose graded components give an exact, algebraic description of multi-hop
reachability. We use this structure to define a sparse, multi-hop attention: the
path algebra supplies the attention \emph{support} (which vertices a node reaches
by a length-$k$ walk) and the network learns the \emph{weights}. On bottleneck
graphs with tunable over-squashing we prove the separation---message passing
with linear aggregation is at exactly chance accuracy for every depth and
width, while a single Walk Attention layer at the matching range solves the
retrieval task exactly---and confirm it empirically: Walk Attention reaches
$1.000\pm0.000$ accuracy over five seeds where one-hop graph attention degrades
toward chance and a fixed-weight aggregator only partially succeeds; it stays
exact on the standard \textsc{RingTransfer} probe and, under a shared small
training budget, is comparable to one-hop attention on the real-world
Peptides-func benchmark---its advantage is decisive when over-squashing is
severe and modest when it is mild. A negative result motivates the
design: using the path-algebra \emph{quotient} to collapse equivalent walks
discards multiplicity the task needs and fails, so the algebra is valuable not for
compression but as the object that \emph{defines the attention support}.
Conceptually, Walk Attention is the learned, representational reading of an
algebraic substrate whose dynamical reading is the automata-of-impact framework.

\keywords{Graph neural networks \and Over-squashing \and Path algebra \and
Quiver \and Representation theory \and Attention.}
\end{abstract}

% =====================================================================
\section{Introduction}
% =====================================================================

Graph neural networks (GNNs) built on the message-passing paradigm
\cite{gilmer2017neural} update each node by aggregating messages from its
immediate neighbours and iterating this local rule across layers. After $k$
layers, information from nodes $k$ hops away reaches a target node. While simple
and effective, this scheme has a well-documented failure mode:
\emph{over-squashing} \cite{alon2021bottleneck}. When the number of paths
carrying information into a node grows---typically exponentially with
distance---the messages along all of those paths must be compressed into a
single fixed-width vector. The receiving node cannot recover them individually,
and long-range signal is lost. Over-squashing is the principal obstacle to learning tasks whose
labels depend on distant or globally-distributed information.

Existing remedies act on the graph or the aggregation: \emph{rewiring} adds or
reweights edges to widen bottlenecks \cite{topping2022understanding}, graph
\emph{coarsening} and \emph{pooling} reduce the graph while trying to preserve
structure, and curvature-based analyses explain \emph{where} squashing occurs
\cite{digiovanni2023oversquashing}. A complementary line of work replaces local
aggregation by global \emph{attention} as in Transformers
\cite{vaswani2017attention}, at the cost of an all-pairs interaction that
ignores graph structure.

In this paper we take a different route, grounded in the representation theory of
quivers. A \emph{quiver} is a directed graph, possibly with multiple arrows
\cite{schiffler2014quiver,assem2006elements}, and its \emph{path algebra} $\kQ$
has the directed walks of the quiver as a basis, with concatenation as
multiplication. The graded piece $e_j \cdot \kQ_k \cdot e_i$ of this algebra has
a basis consisting of the directed walks of length~$k$ from vertex~$i$ to
vertex~$j$; its dimension equals $(\Adj^k)_{ij}$, the corresponding entry of the
$k$-th power of the adjacency matrix. The path algebra therefore provides an
\emph{exact, algebraic description of multi-hop reachability and multiplicity}.

Our central observation is that this algebraic structure is precisely what is
needed to define a principled, sparse, multi-hop attention. We let each node
attend directly to every vertex reachable from it by a walk of length~$k$: the
path algebra supplies the \emph{support} of the attention (which pairs may
interact at range~$k$), and the network learns the \emph{weights} (how much).
Because the walk-reachability support is sparse---typically a small fraction of
all node pairs---this is far cheaper than full self-attention, while still acting
at a distance and thereby bypassing the iterative bottleneck that causes
over-squashing. We call the resulting architecture \textbf{Walk Attention}.

This places Walk Attention on a common algebraic ground (Section~\ref{sec:ground}):
since a graph network on $Q$ is a representation of the quiver
($\mathrm{Rep}(Q)\cong\kQ\text{-}\mathrm{Mod}$), it is the \emph{learned}
reading of a substrate whose \emph{dynamical} reading is the
automata-of-impact-over-quivers framework \cite{zainea2026aiq}.

\paragraph{Contributions.}
\begin{enumerate}
  \item We give a \emph{self-contained} construction (Section~\ref{sec:algebra})
  of quivers, path algebras, graded components, the impact-degree neighbourhood
  system, and the associated \emph{walk operators}, developing the needed
  representation theory from first principles for a pattern-recognition audience.
  \item We define \textbf{Walk Attention} (Section~\ref{sec:model}): a learned
  multi-hop attention whose support is the path-algebra walk-reachability
  structure, and we relate it to message passing and to Transformer attention.
  \item We \emph{prove} the separation on the bottleneck family
  (Section~\ref{sec:theory}): with the standard one-layer-per-hop depth, any
  message-passing network whose first aggregation is linear in the source
  features (GCN, GIN) has exactly chance accuracy for every depth, width, and
  parameter setting, whereas a single Walk Attention layer at the matching
  range solves the retrieval task exactly.
  \item On a controlled bottleneck family with tunable over-squashing
  (Section~\ref{sec:exp}), Walk Attention attains $1.000\pm0.000$ accuracy over
  five seeds where one-hop graph attention degrades toward chance and a
  fixed-weight multi-hop aggregator only partially succeeds; on the real
  Peptides-func benchmark, under a shared small training budget, it is
  comparable to one-hop attention, so its advantage is regime-dependent
  (decisive under severe over-squashing, modest when it is mild).
  \item We report a \emph{negative} result that sharpens the design: collapsing
  equivalent walks via the path-algebra \emph{quotient} discards needed
  multiplicity and fails. The quotient's role is to \emph{determine} the support, not
  to compress the signal.
\end{enumerate}

% =====================================================================
\section{Related Work}\label{sec:related}
% =====================================================================

\paragraph{Over-squashing.}
Alon and Yahav \cite{alon2021bottleneck} identified over-squashing and proposed
the \textsc{NeighborsMatch} probe, on which standard GNNs degrade sharply with
the problem radius. Topping et al.\ \cite{topping2022understanding} connected
over-squashing to negative graph curvature and proposed stochastic discrete
Ricci flow (SDRF) rewiring; Di Giovanni et al.\
\cite{digiovanni2023oversquashing} gave a sensitivity analysis relating
squashing to width, depth, and topology. These works either modify the graph or
characterise the phenomenon. We instead change the aggregation operator,
attending at a distance over an algebraically-defined support.

\paragraph{Attention and graph Transformers.}
Graph attention networks \cite{velickovic2017graph} learn one-hop attention
weights; full graph Transformers apply Transformer self-attention
\cite{vaswani2017attention} over all node pairs, recovering long range at
quadratic cost and with positional/structural encodings to reinject topology.
Walk Attention is intermediate between the two: it is attention, but its support
is the sparse multi-hop walk-reachability of the quiver rather than one hop or all
pairs.

\paragraph{Multi-hop aggregation and attention.}
Aggregating beyond one hop is well explored: MixHop \cite{abu2019mixhop} and
SIGN \cite{frasca2020sign} mix fixed powers of (normalised) adjacency operators,
graph diffusion convolution \cite{gasteiger2019diffusion} rewires by a diffusion
kernel, MAGNA \cite{wang2021magna} diffuses attention scores over multiple hops,
shortest-path message passing aggregates over $k$-hop distance shells
\cite{abboud2022shortest}, NAGphormer \cite{chen2023nagphormer} tokenises
hop-wise aggregates for a Transformer, Graphormer \cite{ying2021graphormer}
biases dense attention by shortest-path distance, and DRew
\cite{gutteridge2023drew} rewires message passing dynamically with delay. Walk
Attention differs in the object that defines the support: the graded piece of
the path algebra---directed \emph{walks} with their multiplicity, not distance
shells or diffusion weights---which makes the support exact
(Proposition~\ref{prop:count}), exposes the quotient $\kQ/I$ as a design degree
of freedom (Section~\ref{sec:exp}), and embeds the architecture in the
representation theory of quivers (Section~\ref{sec:ground}).

\paragraph{Path and walk structure.}
Path-counting and powers of the adjacency matrix are classical, and the path
algebra of a quiver is a standard object in representation theory
\cite{schiffler2014quiver,assem2006elements}. Brauer-configuration algebras and
related path-algebraic constructions have recently been applied to combinatorial
and applied problems by Moreno Ca\~nadas and collaborators
\cite{moreno2021brauer,moreno2024integer}. Closest to the present work is the
impact-degree neighbourhood system and the automata-of-impact-over-quivers
framework \cite{zainea2026aiq,ibanez2025casimulations}, which we take as the
dynamical sibling of Walk Attention over the same algebraic substrate
(Section~\ref{sec:ground}). Attention over multi-hop supports is thus not new in
itself; what is new, to our knowledge, is taking the support from the
\emph{graded path algebra} of the quiver, with its multiplicities, its quotient,
and its dynamical (automata) reading.

\input{sections/algebra}
\input{sections/model}
\input{sections/experiments}
\input{sections/conclusion}

% =====================================================================
\bibliographystyle{splncs04}
\bibliography{walk_attention}
% =====================================================================

\end{document}
