% =====================================================================
\section{Walk Attention}\label{sec:model}
% =====================================================================

We now define the architecture. The guiding principle is the observation that the
multi-hop walk operator of Section~\ref{sec:algebra} is structurally an
\emph{attention}: it aggregates, for each target node, a weighted combination of
the features of the nodes that reach it. Transformer self-attention
\cite{vaswani2017attention} has the same shape but with all-pairs support and
\emph{learned} weights; graph attention \cite{velickovic2017graph} restricts the
support to one hop. Walk Attention keeps the learned weights of attention and
takes the support from the path algebra: the multi-hop, sparse walk-reachability
$\mathcal{S}_k$. In the terms of Section~\ref{sec:ground}, it is the learned node
update $\varphi$ of the common substrate---a graded representation of $Q$ whose
arrow maps are optimised over the fundamental neighbourhood system.

% ---------------------------------------------------------------------
\subsection{The Walk Attention layer}
% ---------------------------------------------------------------------

Fix a range $k$ with reachability support $\mathcal{S}_k$
(Definition~\ref{def:walkop}). A Walk Attention layer with $H$ heads and head
dimension $C$ transforms node features $x\in\RR^{n\times d_{\text{in}}}$ as
follows. For each head $h$ it computes queries, keys and values by linear maps,
\[
  Q^h = x\,W^h_Q,\qquad K^h = x\,W^h_K,\qquad V^h = x\,W^h_V,
  \qquad W^h_\bullet\in\RR^{d_{\text{in}}\times C},
\]
and then attends \emph{only over the reachable pairs}: for every
$(j,i)\in\mathcal{S}_k$,
\begin{align}
  \ell^h_{ji} &= \frac{1}{\sqrt{C}}\,\langle Q^h_j,\,K^h_i\rangle, \label{eq:logit}\\
  \alpha^h_{ji} &= \softmax_{\,i:\,(j,i)\in\mathcal{S}_k}\bigl(\ell^h_{ji}\bigr),
    \label{eq:softmax}\\
  z^h_j &= \sum_{i:\,(j,i)\in\mathcal{S}_k}\alpha^h_{ji}\,V^h_i. \label{eq:agg}
\end{align}
The softmax in \eqref{eq:softmax} is taken, for each target $j$, over exactly the
sources $i$ that reach $j$ by a length-$k$ walk---a \emph{segmented} softmax over
the support, never over all $n$ nodes. The head outputs are concatenated (or
averaged in the final layer) and a residual ``root'' term is added:
\[
  \mathrm{WA}_k(x)_j \;=\; \bigl\Vert_{h=1}^{H} z^h_j \;+\; x_j W_{\!\text{root}}.
\]
Because \eqref{eq:logit}--\eqref{eq:agg} only ever touch the pairs in
$\mathcal{S}_k$, the layer costs $O(|\mathcal{S}_k|\,H\,C)$ time and memory, not
$O(n^2)$. The support is precomputed once per graph from
Proposition~\ref{prop:count} (the nonzero pattern of $\Adj^{\,k}$).
Figure~\ref{fig:layer} summarises the data flow.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  >={Stealth[round]}, node distance=6mm,
  box/.style={draw,rounded corners=2pt,minimum height=7mm,inner sep=3pt,font=\small,fill=blue!4},
  alg/.style={draw,rounded corners=2pt,minimum height=7mm,inner sep=3pt,font=\small,fill=green!6},
  op/.style={font=\small},
  lbl/.style={font=\scriptsize,gray}]
  \node[box] (x) {$x$};
  \node[box,right=8mm of x] (q) {$Q=xW_Q$};
  \node[box,below=2mm of q] (k) {$K=xW_K$};
  \node[box,below=2mm of k] (v) {$V=xW_V$};
  \node[alg,right=10mm of k] (att)
    {$\displaystyle \alpha_{ji}=\softmax_{i:(j,i)\in\mathcal{S}_k}\!\frac{\langle Q_j,K_i\rangle}{\sqrt C}$};
  \node[box,right=10mm of att] (agg) {$z_j=\sum_i \alpha_{ji}V_i$};
  \node[box,right=8mm of agg] (out) {$\mathrm{WA}_k(x)$};
  % support callout
  \node[alg,below=8mm of att,fill=green!10] (supp)
    {support $\mathcal{S}_k=\{(j,i):(\Adj^{\,k})_{ij}>0\}$};
  \draw[->] (x) -- (q.west);
  \draw[->] (x) |- (k.west);
  \draw[->] (x) |- (v.west);
  \draw[->] (q.east) -- (att.north west);
  \draw[->] (k.east) -- (att.west);
  \draw[->] (att.east) -- (agg.west);
  \draw[->] (v.east) -- ([yshift=-2mm]agg.west);
  \draw[->] (agg) -- (out);
  \draw[->,green!50!black] (supp.north) -- (att.south)
    node[midway,right,lbl,text=green!50!black] {restricts pairs};
  \draw[->,rounded corners] (x.north) -- ++(0,6mm) -| (out.north)
    node[pos=0.25,above,lbl] {root (residual)};
\end{tikzpicture}
\caption{One Walk Attention layer at range $k$. Queries, keys and values are
linear maps of the node features; attention is computed \emph{only} over the
walk-reachable pairs $\mathcal{S}_k$ supplied by the path algebra (the nonzero
pattern of $\Adj^{\,k}$), then aggregated and added to a residual root term. The
green support box is what distinguishes Walk Attention from dense ($\mathcal{S}_k
=$ all pairs) and one-hop ($\mathcal{S}_k = $ edges) attention.}
\label{fig:layer}
\end{figure}

\begin{remark}[Why learned weights matter]
The fixed-weight operator $W^{(k)}$ of Definition~\ref{def:walkop} is the special
case in which $\alpha^h_{ji}$ is replaced by the constant $n_k(i,j)/Z_k$.
It reaches every range-$k$ source but weights them by multiplicity alone, so it
cannot \emph{select} which source carries the relevant signal. The learned
attention \eqref{eq:logit}--\eqref{eq:softmax} restores this selectivity while
keeping the same multi-hop support. Section~\ref{sec:exp} shows this distinction
is decisive.
\end{remark}

% ---------------------------------------------------------------------
\subsection{The network}
% ---------------------------------------------------------------------

A depth-$L$ Walk Attention network stacks one layer per range
$k=1,\dots,L$. Layer $k$ uses the support $\mathcal{S}_k$ derived from
$\Adj^{\,k}$, so successive layers read from progressively longer walks. Each
layer is followed by layer normalisation, an \textsc{elu} nonlinearity and
dropout; a final multilayer perceptron maps the node embedding to task outputs:
\[
  x^{(k)} \;=\; \mathrm{dropout}\Bigl(\mathrm{elu}\bigl(\mathrm{LN}\bigl(
  \mathrm{WA}_k\bigl(x^{(k-1)}\bigr)\bigr)\bigr)\Bigr),\qquad
  \hat y \;=\; \mathrm{MLP}\bigl(x^{(L)}\bigr).
\]
Layer normalisation is important in practice: without it the learned attention is
high-variance across initialisations (it reaches the optimum on some seeds and
not others); with it, together with a short learning-rate warmup and gradient
clipping, training converges reliably (Section~\ref{sec:exp}).
Algorithm~\ref{alg:wa} summarises one layer.

\begin{algorithm}[t]
\caption{Walk Attention layer at range $k$ (sparse).}\label{alg:wa}
\begin{algorithmic}[1]
\Require features $x\in\RR^{n\times d_{\text{in}}}$; support
  $\mathcal{S}_k=\{(j,i):(\Adj^{\,k})_{ij}>0\}$
\State $Q,K,V \gets xW_Q,\,xW_K,\,xW_V$ \Comment{per head; shapes $n\times HC$}
\For{each $(j,i)\in\mathcal{S}_k$ and head $h$}
  \State $\ell^h_{ji}\gets \langle Q^h_j,K^h_i\rangle/\sqrt{C}$
\EndFor
\State $\alpha^h_{ji}\gets$ segmented-$\softmax$ over $\{i:(j,i)\in\mathcal{S}_k\}$
\State $z^h_j\gets \sum_i \alpha^h_{ji}V^h_i$
  \Comment{scatter-add into targets}
\State \Return $\bigl\Vert_h z^h \;+\; xW_{\!\text{root}}$
\end{algorithmic}
\end{algorithm}

% ---------------------------------------------------------------------
\subsection{Relation to message passing and Transformers}
% ---------------------------------------------------------------------

Table~\ref{tab:relation} situates Walk Attention among related operators by two
axes: the \emph{support} of interaction and whether the weights are
\emph{learned}. A one-hop graph-attention layer \cite{velickovic2017graph} learns
weights but only over direct neighbours, so it must propagate long-range signal
through $L$ stacked layers---precisely the regime where over-squashing is most
severe. A full graph Transformer \cite{vaswani2017attention} learns weights over
all pairs,
paying $O(n^2)$ and discarding graph structure unless it is reinjected by
positional encodings. The fixed-weight walk operator has the right multi-hop
support but cannot select. Walk Attention occupies the remaining cell: multi-hop,
structure-aware support \emph{and} learned, selective weights.

\begin{table}[t]
\centering
\caption{Walk Attention against related aggregation operators.}
\label{tab:relation}
\begin{tabular}{lccc}
\toprule
\textbf{Operator} & \textbf{Support} & \textbf{Weights} & \textbf{Cost} \\
\midrule
GCN / message passing       & 1-hop      & fixed (degree) & $O(|Q_1|)$ \\
Graph attention (GAT)       & 1-hop      & learned        & $O(|Q_1|)$ \\
Walk operator $W^{(k)}$     & walk-reachable & fixed (multiplicity) & $O(|\mathcal{S}_k|)$ \\
Graph Transformer           & all pairs  & learned        & $O(n^2)$ \\
\textbf{Walk Attention (ours)} & \textbf{walk-reachable} & \textbf{learned} & $O(|\mathcal{S}_k|)$ \\
\bottomrule
\end{tabular}
\end{table}

% ---------------------------------------------------------------------
\subsection{A formal account on the bottleneck family}\label{sec:theory}
% ---------------------------------------------------------------------

On the bottleneck-chain family used in Section~\ref{sec:exp} the separation
between one-hop message passing and Walk Attention can be \emph{proved}, not
only measured. Recall the construction: $K$ source vertices carry a one-hot
\emph{key}---the assignment of keys to sources is a uniformly random
permutation $\pi$---and a one-hot \emph{value}; $d$ bottleneck layers
$L_1,\dots,L_d$ of $M$ interchangeable, featureless vertices follow,
consecutive layers fully connected; a target vertex carries a one-hot
\emph{query} $q$ and must output the value of the source whose key equals $q$
(label $\pi^{-1}(q)$, chance $1/K$). Throughout, networks have the standard
depth $L=d+1$---one layer per hop, as in the \textsc{NeighborsMatch} protocol
\cite{alon2021bottleneck} and in our experiments---and hidden width $w$.

\begin{proposition}[Layer collapse]\label{prop:collapse}
For any message-passing network on this graph whose update
$h_v' = U\bigl(h_v,\mathrm{AGG}\{h_u : u\to v\}\bigr)$ uses the same $U$ and
$\mathrm{AGG}$ at every vertex: after every round, all $M$ vertices of each
bottleneck layer carry identical states. Consequently the entire influence of
the sources on the target passes through a single $w$-dimensional state per
layer---while the number of length-$(d{+}1)$ walks carrying it is $K M^{d}$
(Proposition~\ref{prop:count}).
\end{proposition}

\begin{proof}
Induction on the round. Vertices of a bottleneck layer start with identical
(zero) features, and two vertices of the same layer have the same in-multiset
of neighbours (all sources for $L_1$; all of $L_{r-1}$ for $L_r$), whose states
are equal by induction; shared $U,\mathrm{AGG}$ produce equal updates. \qed
\end{proof}

\begin{proposition}[Linear first aggregation is permutation-blind]
\label{prop:blind}
Assume in addition that messages are linear in the sender's state with
coefficients independent of feature content---as in GCN (degree-normalised
linear messages) and GIN (sum of neighbour states). Then with $L=d+1$ layers
the target's output depends on the sources only through $\sum_{s}x_s$, which is
the \emph{same for every} $\pi$: the key slots of a permutation sum to the
all-ones vector. The expected accuracy is therefore exactly $1/K$, for every
depth $d$, width $w$, and choice of parameters.
\end{proposition}

\begin{proof}
In-neighbours of the target are $L_d$ (and itself); unrolling, the target's
state after round $d{+}1$ depends on the sources only through the state of
$L_1$ after round~$1$: source information entering $L_1$ at round $t$ needs $d$
further hops to reach the target, arriving at round $t+d\le d+1$ only if
$t\le1$. At round $1$ the message from source $s$ is a fixed linear map of the
raw feature $x_s$, with one shared coefficient (all sources are isomorphic in
the graph), so the state of $L_1$ is a function of $\sum_s x_s$. With keys
$e_{\pi(s)}$, the key block of $\sum_s x_s$ equals $\sum_s e_{\pi(s)}
=(1,\dots,1)$ for every $\pi$, and the value block is likewise
$\pi$-independent; hence the output is independent of $\pi$. Given the query
$q$, the label $\pi^{-1}(q)$ is uniform on $\{1,\dots,K\}$, so any
$\pi$-independent prediction has expected accuracy $1/K$. \qed
\end{proof}

\begin{proposition}[One Walk Attention layer suffices]\label{prop:wa}
At range $k=d{+}1$ the support of the target is exactly the set of $K$ sources
(the graph is graded; Proposition~\ref{prop:fns}). Choose $W_Q$ to project the
query slot scaled by any $\beta>0$, $W_K$ the key slot, $W_V$ the value slot,
and $W_{\!\text{root}}=0$. Then the Walk Attention output at the target is
$z_t=\sum_s \alpha_s\,e_{v(s)}$ whose largest entry sits at the coordinate of
the correct value, on \emph{every} instance and for all $d$ and $M$.
\end{proposition}

\begin{proof}
The logit of source $s$ is $\beta\langle e_q, e_{\pi(s)}\rangle
=\beta\,\mathbf{1}[\pi(s)=q]$, so the softmax over the $K$ reachable sources
puts weight $e^{\beta}/(e^{\beta}+K-1)$ on the unique matching source $s^\ast$
and equal smaller weight on the rest. The values $e_{v(s)}$ are distinct basis
vectors, so the largest coordinate of $z_t$ is $\alpha_{s^\ast}$, at the
coordinate $v(s^\ast)=\pi^{-1}(q)$, the label. \qed
\end{proof}

\begin{remark}\label{rem:mechanisms}
The propositions delimit the mechanisms seen in Section~\ref{sec:exp}. GAT
escapes Proposition~\ref{prop:blind}---its first-round coefficients depend on
source content---but must still thread the key--value association through the
single-vector channel of Proposition~\ref{prop:collapse}, re-encoded $d{+}1$
times: empirically it sits above chance and degrades with depth. The
fixed-weight walk operator reads the sources directly but with \emph{equal}
weights (all multiplicities equal $M^{d}$), so it must embed the whole
key--value multiset in $w$ dimensions rather than select one source: it
plateaus in between. Walk Attention both reads directly and selects
(Proposition~\ref{prop:wa}), and solves the task.
\end{remark}
